Clines from PDE

### 2015-12-11 notes/pde-clines.Rmd
fixed symmetry bug?

First, we compute per-allele solutions:
```{r solve_LD, cache=TRUE}
s <- 0.1
tt <- seq(0,100,length.out=21)
xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
rr <- seq(0,0.5,length.out=20)[-1]
fwds.soln <- forwards_pde(s=s,times=tt,grid=xgrid)
LD.solns <- lapply( rr, function (r) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=c(0,r),N=2), fwds.soln=fwds.soln) )
```
... but then we must extract linkage from these.
Below we report for recombination distance $r$,
the probability that two markers at distance $r$ 
sampled in one individual at that location and time
(with one end fixed to the selected site)
are both of the same ancestral type.
If the local proportion of type $A$ is $p$ and the probability when linked to $A$ is $u_A$,
we want $p u_A(r_0,r_0+r) + (1-p) u_B(r_0,r_0+r) - (p u_A(r_0) + (1-p) u_A(r_0))(p u_A(r_0+r) + (1-p) u_A(r_0+r))$:
```{r compute_LD, cache=TRUE, depends="solve_LD"}
pfun <- function (t,x) { 
        fp <- function (x,t) { approx(xgrid$x.mid,cline_interp(t,soln=fwds.soln),xout=x)$y }
        ans <- numeric(length(x))
        for (tval in unique(t)) {
            dothese <- (rep_len(t,length(x))==tval)
            ans[dothese] <- fp(x=x[dothese],t=tval)
        }
        return(ans)
}
ABfac <- rep( rep(c("A","B"),each=xgrid$N), 3 )  # three combinations of two recomb points
rfac <- rep( c("11","12","22"), each=xgrid$N*2 )
LDs <- lapply( LD.solns, function (lds) {
            # returns a time x space matrix
            pp <- outer( lds[,1], xgrid$x.mid, pfun )
            return( 
                   pp * lds[,1+which(ABfac=="A"&rfac=="12")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="12")] 
                   - ( pp * lds[,1+which(ABfac=="A"&rfac=="11")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="11")] )
                       * ( pp * lds[,1+which(ABfac=="A"&rfac=="22")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="22")] )
               )
    } )
plot.x <- xgrid$x.mid[which.min(abs(xgrid$x.mid))]
matplot( tt, sapply( LDs, function (x) { x[,match(plot.x,xgrid$x.mid)] } ), type='l', ylab="ancestral LD", xlab="time", main=paste("location=",round(plot.x,2)) )
plot.t <- tt[which.min(abs(tt-20))]
matplot( xgrid$x.mid, sapply( LDs, function (x) { x[match(plot.t,tt),] } ), type='l', ylab="ancestral LD", xlab="space" )
matplot( rr, t(sapply( LDs, function (x) { x[match(plot.t,tt),] } )), type='l', ylab="ancestral LD", xlab="recomb dist", main=paste("t=",round(plot.t,2)), col=rainbow(xgrid$N), lty=1 )
legend("topright",lty=1,col=rainbow(xgrid$N)[floor(seq(1,xgrid$N,length.out=6))],legend=paste("x=",round(xgrid$x.mid[floor(seq(1,xgrid$N,length.out=6))],2)))
```
Hm, the right/left spatial positions should look the same, right?  Is something off?

>>>>

First, we compute per-allele solutions:
```{r solve_LD, cache=TRUE}
s <- 0.1
tt <- seq(0,100,length.out=21)
xgrid <- setup.grid.1D(x.up=-20.5, x.down=20.5, N=41)
rr <- seq(0,0.5,length.out=20)[-1]
fwds.soln <- forwards_pde(s=s,times=tt,grid=extend_grid(xgrid))
LD.solns <- lapply( rr, function (r) forwards_backwards_haplotypes(s=s, times=tt, xgrid=xgrid, rgrid=list(x.mid=c(0,r),N=2), fwds.soln=fwds.soln) )
```
... but then we must extract linkage from these.
Below we report for recombination distance $r$,
the probability that two markers at distance $r$ 
sampled in one individual at that location and time
(with one end fixed to the selected site)
are both of the same ancestral type.
If the local proportion of type $A$ is $p$ and the probability when linked to $A$ is $u_A$,
we want $p u_A(r_0,r_0+r) + (1-p) u_B(r_0,r_0+r) - (p u_A(r_0) + (1-p) u_A(r_0))(p u_A(r_0+r) + (1-p) u_A(r_0+r))$:
```{r compute_LD, cache=TRUE, depends="solve_LD"}
pfun <- function (t,x) { 
        fp <- function (x,t) { approx(xgrid$x.mid,cline_interp(t,soln=fwds.soln),xout=x)$y }
        ans <- numeric(length(x))
        for (tval in unique(t)) {
            dothese <- (rep_len(t,length(x))==tval)
            ans[dothese] <- fp(x=x[dothese],t=tval)
        }
        return(ans)
}
ABfac <- rep( rep(c("A","B"),each=xgrid$N), 3 )  # three combinations of two recomb points
rfac <- rep( c("11","12","22"), each=xgrid$N*2 )
LDs <- lapply( LD.solns, function (lds) {
            # returns a time x space matrix
            pp <- outer( lds[,1], xgrid$x.mid, pfun )
            return( 
                   pp * lds[,1+which(ABfac=="A"&rfac=="12")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="12")] 
                   - ( pp * lds[,1+which(ABfac=="A"&rfac=="11")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="11")] )
                       * ( pp * lds[,1+which(ABfac=="A"&rfac=="22")] + (1-pp) * lds[,1+which(ABfac=="B"&rfac=="22")] )
               )
    } )
plot.x <- xgrid$x.mid[which.min(abs(xgrid$x.mid))]
matplot( tt, sapply( LDs, function (x) { x[,match(plot.x,xgrid$x.mid)] } ), type='l', ylab="ancestral LD", xlab="time", main=paste("location=",round(plot.x,2)) )
plot.t <- tt[which.min(abs(tt-20))]
matplot( xgrid$x.mid, sapply( LDs, function (x) { x[match(plot.t,tt),] } ), type='l', ylab="ancestral LD", xlab="space" )
matplot( rr, t(sapply( LDs, function (x) { x[match(plot.t,tt),] } )), type='l', ylab="ancestral LD", xlab="recomb dist", main=paste("t=",round(plot.t,2)), col=rainbow(xgrid$N), lty=1 )
legend("topright",lty=1,col=rainbow(xgrid$N)[floor(seq(1,xgrid$N,length.out=6))],legend=paste("x=",round(xgrid$x.mid[floor(seq(1,xgrid$N,length.out=6))],2)))
```
Hm, the right/left spatial positions should look the same, right?  Is something off?
